#Reproducible Research: Peer Assessment 1
##Introduction
It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.
This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

##Data

The data for this assignment has been downloaded from the course web site [ Activity monitoring data] [1] and stored locally as __activity.csv__
The variables included in this dataset are:
* __steps__: Number of steps taking in a 5-minute interval (missing values are
coded as NA)
* __date__: The date on which the measurement was taken in YYYY-MM-DD
format
* __interval__: Identifier for the 5-minute interval in which measurement was taken
The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

##Loading and preprocessing the data
Show any code that is needed to
* Load the data (i.e. read.csv())
* Process/transform the data (if necessary) into a format suitable for your
analysis

```{r loading, echo=TRUE}
data <- read.csv("activity.csv")
data$date <- as.Date(data$date)
```

## What is mean total number of steps taken per day?
For this part of the assignment, __you can ignore the missing values__ in the dataset.
* Make a histogram of the total number of steps taken each day
* Calculate and report the mean and median total number of steps taken per day

```{r number of steps, echo=TRUE}
hist(tapply(data$steps,data$date,sum,na.rm=TRUE),main="Histogram - Total number of steps taken each day",col="red",xlab="Number of steps per day")
mean(tapply(data$steps,data$date,sum,na.rm=TRUE))
median(tapply(data$steps,data$date,sum,na.rm=TRUE))
```

__The mean total number of steps taken per day is `r mean(tapply(data$steps,data$date,sum,na.rm=TRUE))` steps per day and the median is `r median(tapply(data$steps,data$date,sum,na.rm=TRUE))` steps per day__

## What is the average daily activity pattern?
* Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r Series plot, echo=TRUE}
MedSteps <- tapply(data$steps,data$interval,mean,na.rm=TRUE)
plot(dimnames(MedSteps)[[1]],MedSteps,type="l",main="Average number of steps in interval of 5 minutes",xlab="Minute of the day",ylab="Steps per 5 minute")
```

* Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r max_steps, echo=TRUE}
intervalo = which.max(MedSteps)
intervalo
```
__The 5-minute interval, on average across all the days in the dataset, which contains the maximum number or steps is the interval `r names(intervalo)`__


## Imputing missing values
Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some
calculations or summaries of the data.
* Calculate and report the total number of missing values in the dataset
(i.e. the total number of rows with NAs)

Examining the whole dataset, we can see the general characteristics of the data
```{r MissingValues,echo=TRUE}
summary(data)
```
As it can be seen in the summary, __there is only NA's in the steps field__

* Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
```{r strategyFilling,echo=TRUE}
# Analizing where are the NA
table(data$date,is.na(data$step))
# There are days with NA all day --> We can't use the mean/median for that day
table(data$interval,is.na(data$step))
# There is no interval with NA all interval --> We can use the mean/media for each interval as a strategy for filling the NA
```

* Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r Filling na, echo=TRUE}
# Mean step across all days --> MedSteps
MedSteps <- tapply(data$steps,data$interval,mean,na.rm=TRUE)
# Function to filling na with mean from MedSteps
filling <- function (data) {
    temp <- data
    for (i in 1:nrow(data)){
        if (is.na(data$steps[i])==TRUE){
            temp$steps[i]=MedSteps[as.character(data$interval[i])]
        }
    }
    temp
}
# New data set without NA
datafilled = filling (data)
summary(datafilled)
```

* Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r new number of steps, echo=TRUE}
# Histogram with new data without NA
hist(tapply(datafilled$steps,datafilled$date,sum),main="Histogram 2- Total number of steps taken each day (filled NA)",col="red",xlab="Number of steps per day")
mean(tapply(datafilled$steps,datafilled$date,sum))
median(tapply(datafilled$steps,datafilled$date,sum))
```

### Comparative with and without NA
__Histograms__
```{r comparativeNA, echo=TRUE}
par(mfrow = c(2,1))
hist(tapply(data$steps,data$date,sum,na.rm=TRUE),main="Histogram 1 - Total number of steps taken each day",col="red",xlab="Number of steps per day")
hist(tapply(datafilled$steps,datafilled$date,sum),main="Histogram 2- Total number of steps taken each day (filled NA)",col="red",xlab="Number of steps per day")
```
__Change in mean and median__
```{r comparativeNAstatistics,echo=TRUE}
#With NA
mean(tapply(data$steps,data$date,sum,na.rm=TRUE))
median(tapply(data$steps,data$date,sum,na.rm=TRUE))
#Without NA
mean(tapply(datafilled$steps,datafilled$date,sum))
median(tapply(datafilled$steps,datafilled$date,sum))
```

## Are there differences in activity patterns between weekdays and weekends?
For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.
* Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r filling weekdays and weekend,echo=TRUE}
# New variable in dataset with value "weekday"
datafilled$day="weekday"
# Change to "weekend" if date is "sábado" or "domingo" (locale Spain)
datafilled$day[weekdays(datafilled$date)=="sábado"]="weekend"
datafilled$day[weekdays(datafilled$date)=="domingo"]="weekend"
datafilled$day = as.factor(datafilled$day)
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r graph lattice, echo=TRUE}
library(lattice)

```



[1]: https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip "Activity Monitoring Data"